A New Metric for the Analysis of Swarms using Potential Fields

### 2018-10-18 paper.tex

\begin{abstract}
  % An important criterion in the assessment of swarm coordination algorithms is
  % the extent to which they cause unnecessary movement, i.e. movement which
  % occurs as a side effect of the algorithm rather than from the fulfilment of
  % the intended function of the swarm. 
  There are many metrics defined for the analysis of swarm coordination
  algorithms. These metrics are usually based upon the distances between
  agents, the distance between agents and a fixed point, or the resultant vectors
  that \textit{potential field} effects produce. This paper examines a
  distance-based metric that measures a swarm's overall structure using
  inter-agent distances. It introduces a new metric that identifies a swarm's
  state based upon the resultant magnitude of the vectors produced by the
  agent interactions that create the agent distribution within the swarm's
  structure. The algorithms used to implement the swarming feature are based
  upon cohesion and repulsion vectors between an agent and its neighbours. In
  comparing and contrasting the two metrics, we find that the
  \textit{cohesion/repulsion} metric offers a number of advantages over the
  distance metric. In particular, the \textit{cohesion/repulsion} metric allows
  the identification of the essential characteristic of a swarm as `expanding',
  `stable', or `contracting'. These states cannot be identified using a
  distance-based metric. Practical swarming applications where the new metric
  can be applied advantageously include area-filling and reconnaissance.
\end{abstract}

@@

A swarm can also have a \emph{goal}. This is modelled simply by the addition of 
a \textit{directional vector}, as shown in~Equation~\ref{eq:InterAgentMovementDirectional1}.
\begin{equation}\label{eq:InterAgentMovementDirectional1}
v(b) = k_cv_{c}(b) + k_rv_{r}(b) + k_dv_{d}(b)
\end{equation}
Here, $k_dv_{d}(b)$ is the
calculated directional vector to be applied to an agent.

@@

\begin{center}
\begin{equation}
\label{eq:AgentMovement}
b_{pos}=s_{b}t\big(v(b)\big)\string^ 
\end{equation}
\end{center}

@@

Navarro and Fernando describe a \textit{mean distance error} metric that is
based on the variations in distances between inter-agent spaces~\cite{NIM:09}.
This is the same as the standard deviation of the distance-based internal
movement metric.

@@

% \section{Calculating distance internal movement}
Equation~\ref{eq:SwarmStabilityDistance1} calculates the mean distance of an agent to its neighbours $nbr(b)$. 
%% The relative position vector generated for an agent $b$ to its neighbour $b'$, $bb'$, is shown in~(\ref{eq:FlyToCentre1}). The magnitude of that vector gives the distance between two agents. For an individual agent the average magnitude $\mu_d(b)$ is calculated as \ref{eq:SwarmStabilityDistance1} where $b$ is the agent and $|nbr(b)|$ is the number of neighbours.

@@

The mean distance for a swarm is calculated
by~Equation~\ref{eq:SwarmStabilityDistance2}. All the inter-agent distances are
included for the swarm ($S$).

@@

%\section{Standard deviation in distance metric}\label{Section:VarianceInDistance}
The mean distance provides a indication of the large scale structure of the
swarm. However it is not sufficient to give an indication of the internal
distribution of the agents. The standard deviation clarifies the distribution
within the swarm as shown in~Equation~\ref{eq:SwarmStabilityDistance3}.

@@

@@

This paper defines the cohesion/repulsion metric scalar for an agent as the
length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is,
the weighted effects of cohesion and repulsion by neighbours of the agent. This
number carries a negative sign if repulsion dominates and a positive sign if
cohesion dominates. This is summarised in Equation~\ref{eq:CohesionRepulsion}.

If an agent's {cohesion/repulsion metric} is negative the swarm's bias is to
expand as the repulsion is greater than cohesion. This is seen in the
disorganised stage of a swarm~(Fig.~\ref{coord:BaselineMagnitude1}). The
disorganised stage is the initial random deployment of the agents. If the agent
\textit{agent resultant magnitude} is positive then the swarm is exhibiting a
tendency to contract and this indicates the swarm is a cohesive entity. This
could also be described as the swarm being `sticky' as the agents bias is to
`pull' towards each other.

The \textit{cohesion/repulsion metric} on its own does not give a complete
measure of a swarm's internal state. There needs to be a qualifying component
to the metric that identifies the degree of deviation in the resultant
magnitude, this is the \textit{jitter}. The smaller the deviation the more
uniform the structure of the swarm. These two components identify the degree to
which a swarm has progressed towards a stable state.

The \textit{cohesion/repulsion metric} provides a view of the swarm's state
through the balance between the repulsive and the cohesive vectors that are
being applied to each agent. The deviation component identifies the degree to
which the swarm has stabilised. The ideal status for inter-agent interactions
would be for the agents to have a (\textit{cohesion/repulsion metric}) of zero
or above. This would indicate that the agents are distributed such that they
are at their distribution limit (outer most range of the cohesion field) or at
a level that causes the agents to `pull' together. The ideal degree of
deviation is zero as this indicates an even distribution of agents. These two
aspects of a swarm's features are not considered by Gazi and
Passino~\cite{GP:11} or Barnes et al~\cite{BFV:07} as a means of quantifying
the structure of a swarm in terms of stability.

@@

Section~\ref{Section:SwarmModelling}~discussed calculation of resultant
cohesion~(Equation~\ref{eq:FlyToCentre1}) and resultant
repulsion~(Equation~\ref{eq:Repulsion1}) vectors for each agent deriving a
combined \textit{cohesion/repulsion vector} using as shown in
Equation~\ref{eq:InterAgentMovement2}.

@@

Although it is possible for $v(b)$ to be null there could still be variation in
the constituent components. The variation calculation (standard deviation) as
shown in Equation~\ref{eq:SwarmStabilityQuotientT}.

@@

The mean cohesion/repulsion scalar for the swarm is now given
by~(\ref{eq:SwarmStabilityMetricT3}).

@@

Figure~\ref{coord:BaselineMagnitude1} shows the evolution of the metric in time.

@@

Area filling is a technique used to fill an enclosed area with agents such that
they are distributed as `effectively' as possible throughout an area.

@@

To test this hypothesis a swarm is modelled in the simulator. The model
consists of an obstacle-based enclosed space and a swarm consisting of 60
agents. The experiments parameters for the simulation are shown
in~Table~\ref{tab:FillParameters1}

@@

The field effects are incremented in turn, cohesion range followed by repulsion
range. After each repulsion field change the swarm is allowed to redistribute
itself and stabilise. Table~\ref{tab:FillSequence} shows the field effect
settings that are used for the simulation. The field effects are selected so as
to ensure the swarm parameters have the potential to create a hexagonal swarm.
Figure \ref{emerge:FLOOD0} shows the initial swarm deployment.

@@

Figure~\ref{emerge:FLOOD1} shows the point a which data logging starts. The
swarm at this stage has expanded into the space but is not filling it. This
presents itself as a positive cohesion/repulsion metric value which indicates
the swarm is cohesive and the agents are still ``pulling" together.
Fig.~\ref{emerge:FLOOD6} shows the final swarm distribution for the simulation,
at this point the cohesion/repulsion metric has fell below zero indicating the
swarm is trying to expand.

@@

Figure~\ref{emerge:FIELDFILL-MAG} shows the cohesion/repulsion metric for the
swarm for the entire simulation. Between 100 seconds and 105 seconds there is a
significant change in the magnitude where the value becomes negative. This
indicates there is a compression effect on the swarm which indicates the area
is filled.

@@

Figure~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the same
simulation. The initial deployment is shown at 0 seconds followed by
incremental changes in the distances and variance of the distances as each
increment is made in the field effects. The simulation terminates after 120
seconds. There is no feature of this graph which indicates that the area is
filled. The cohesion/repulsion metric clearly tells us more.

@@

The new \textit{cohesion/repulsion} metric is based upon the algorithms that
have produced the structure and is therefore an analysis of how the algorithms
are performing rather than the structure presented. The new metric also
provides an indication of a swarm's potential future movement. This is
independent of the physical distribution of the agents within the swarm. The
metric also provides the same information as the distance metric in terms of
agent distribution when the field effects are consistent across all agents i.e.
monolithic field effects. This is shown in Fig. \ref{emerge:FIELDFILL-MAG} and
\ref{emerge:FIELDFILL-DIST} by the profiles of the two graphs being the same.
In addition the new metric provides a mechanism for evaluating swarms that use
different algorithms to create a swarming effect and allows the algorithms
inter-agent interactions to be compared rather than the resultant structure the
algorithm creates.

>>>>

\begin{abstract}
  % An important criterion in the assessment of swarm coordination algorithms is
  % the extent to which they cause unnecessary movement, i.e. movement which
  % occurs as a side effect of the algorithm rather than from the fulfilment of
  % the intended function of the swarm. 
  There are many metrics defined for the analysis of swarm coordination
  algorithms. These metrics are usually based upon the distances between
  agents, the distance between agents and a fixed point, or the resultant
  vectors that \textit{potential field} effects produce. This paper examines a
  distance-based metric that measures a swarm's overall structure using
  inter-agent distances. More importantly, it introduces a new metric that
  identifies a swarm's state based upon the resultant magnitude of the vectors
  produced by the agent interactions that create the agent distribution within
  the swarm's structure. The algorithms used to implement the swarming feature
  are based upon cohesion and repulsion vectors between an agent and its
  neighbours. In comparing and contrasting the two metrics, we find that the
  \textit{cohesion/repulsion} metric offers a number of advantages over the
  distance metric. In particular, the \textit{cohesion/repulsion} metric allows
  the identification of the essential characteristic of a swarm as `expanding',
  `stable', or `contracting'. These states cannot be identified using a
  distance-based metric. Practical swarming applications where the new metric
  can be applied advantageously include area-filling and reconnaissance.
\end{abstract}

@@

A swarm can also have a \textit{goal}. This is modelled simply by the addition of 
a \textit{directional vector}, as shown in~Equation~\ref{eq:InterAgentMovementDirectional1}.
\begin{equation}\label{eq:InterAgentMovementDirectional1}
v(b) = k_cv_{c}(b) + k_rv_{r}(b) + k_dv_{d}(b)
\end{equation}
Here, $k_dv_{d}(b)$ is the
calculated directional vector to be applied to an agent.

@@

\begin{center}
\begin{equation}
\label{eq:AgentMovement}
  b_{pos}=s_{b}t\widehat{v(b)} 
\end{equation}
\end{center}

@@

Navarro and Fernando describe a \textit{mean distance error} metric that is
based on the variations in distances between agents~\cite{NIM:09}.
This is the same as the standard deviation of the distance-based internal
movement metric described in this section.

@@

% \section{Calculating distance internal movement}
Equation~\ref{eq:SwarmStabilityDistance1} calculates the mean distance of an agent $b$ to its neighbours $nbr(b)$. 
%% The relative position vector generated for an agent $b$ to its neighbour $b'$, $bb'$, is shown in~(\ref{eq:FlyToCentre1}). The magnitude of that vector gives the distance between two agents. For an individual agent the average magnitude $\mu_d(b)$ is calculated as \ref{eq:SwarmStabilityDistance1} where $b$ is the agent and $|nbr(b)|$ is the number of neighbours.

@@

The mean distance for a swarm is given 
by~Equation~\ref{eq:SwarmStabilityDistance2}. All the inter-agent distances are
included for the swarm ($S$).

@@

%\section{Standard deviation in distance metric}\label{Section:VarianceInDistance}
The mean distance provides an indication of the large scale structure of the
swarm. However it is not sufficient to give an indication of the internal
distribution of the agents. The standard deviation clarifies the distribution
within the swarm as shown in~Equation~\ref{eq:SwarmStabilityDistance3}.

@@

Figure~\ref{coord:BaselineDistance1} shows a short example of the evolution 
of the metric over time.

@@

This paper defines the \textit{cohesion/repulsion metric} for an agent as the
length of $v(b)$,  as defined in Equation~\ref{eq:InterAgentMovement2}: that is
the size of the weighted effects of cohesion and repulsion caused by neighbours of the
agent. This number carries a negative sign if repulsion dominates and a
positive sign if cohesion dominates. This is summarised in
Equation~\ref{eq:CohesionRepulsion}.

If the overall cohesion/repulsion metric of the swarm is negative, the
swarm's tendency is to expand, as the repulsion is greater than cohesion. This is
seen in the disorganised stage of a
swarm~(Fig.~\ref{coord:BaselineMagnitude1}). The disorganised stage is the
initial random deployment of the agents. If the overall cohesion/repulsion
metric is positive, the swarm's tendency is to contract, and
suggests that the swarm is a cohesive entity. This could also be described as
the swarm being `sticky', as the agents' tendency is to `pull' towards each
other.

The cohesion/repulsion metric on its own does not give a complete
measure of a swarm's internal state. There needs to be a qualifying component
to the metric that identifies the degree of deviation in the resultant
magnitudes; this is the \textit{jitter}. The smaller the deviation, the more
uniform the structure of the swarm. These two components identify the degree to
which a swarm has progressed towards a stable state.

The cohesion/repulsion metric provides a view of the swarm's state
through the balance between the repulsive and the cohesive vectors that are
being applied to each agent. The deviation component identifies the degree to
which the swarm has stabilised. The ideal status for inter-agent interactions
would be for the agents to have a cohesion/repulsion metric of zero
or above. This would indicate that the agents are distributed such that they
are at their distribution limit (outer most range of the cohesion field) or at
a level that causes the agents to `pull' together. The ideal degree of
deviation is zero as this indicates an even distribution of agents. These two
aspects of a swarm's features are not considered by Gazi and
Passino~\cite{GP:11} or Barnes et al~\cite{BFV:07} as a means of quantifying
the structure of a swarm in terms of stability.

@@

Section~\ref{Section:SwarmModelling}~discussed calculation of resultant
cohesion~(Equation~\ref{eq:FlyToCentre1}) and resultant
repulsion~(Equation~\ref{eq:Repulsion1}) vectors for each agent, deriving a
combined cohesion/repulsion vector, as shown in
Equation~\ref{eq:InterAgentMovement2}.

@@

Although it is possible for $v(b)$ to be null there could still be variation in
the constituent components. The variation calculation (standard deviation) is
shown in Equation~\ref{eq:SwarmStabilityQuotientT}.

@@

The mean cohesion/repulsion scalar for the swarm is now given
by~Equation~\ref{eq:SwarmStabilityMetricT3}.

@@

Figure~\ref{coord:BaselineMagnitude1} shows a short example of the evolution of
the metric over time.

@@

Area filling is a technique used to fill an enclosed area with agents such that
they are distributed as `effectively' as possible throughout the area.

@@

To test this hypothesis a swarm is modelled in the simulator. The model
consists of an obstacle-based enclosed space and a swarm consisting of 60
agents. The experimental parameters for the simulation are shown
in~Table~\ref{tab:FillParameters1}

@@

The field effects are incremented in turn, cohesion range followed by repulsion
range. After each pair of increments, the swarm is allowed to redistribute
itself and stabilise. Table~\ref{tab:FillSequence} shows the field effect
settings that are used for the simulation. The field effects are selected so as
to ensure the swarm parameters have the potential to create a hexagonal swarm.
Figure \ref{emerge:FLOOD0} shows the initial swarm deployment.

@@

Figure~\ref{emerge:FLOOD1} shows the point at which data logging starts. The
swarm at this stage has expanded into the space but is not filling it. This
presents itself as a positive cohesion/repulsion metric value, which indicates
the swarm is cohesive and the agents are still ``pulling" together.
Fig.~\ref{emerge:FLOOD6} shows the final swarm distribution for the simulation.
At this point the cohesion/repulsion metric has fallen below zero, indicating
the swarm is trying to expand.

@@

Figure~\ref{emerge:FIELDFILL-MAG} shows the cohesion/repulsion metric for the
swarm for the entire simulation. Between 100 seconds and 105 seconds there is a
significant change in the magnitude where the value becomes negative. This
indicates that there is a compression effect on the swarm and, hence, that the
area has been filled.

@@

Figure~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the same
simulation. The initial deployment is shown at 0 seconds, followed by
incremental changes in the distances, and variance of the distances, as each
increment is made in the field effects. The simulation terminates after 120
seconds. There is no feature of this graph which indicates that the area is
filled. The cohesion/repulsion metric clearly tells us more.

@@

The new \textit{cohesion/repulsion} metric is based upon the algorithms that
have produced the structure and is therefore an analysis of how the algorithms
are performing rather than the structure presented. The new metric also
provides an indication of a swarm's potential future movement. This is
independent of the physical distribution of the agents within the swarm. The
metric also provides the same information as the distance metric in terms of
agent distribution when the field effects are consistent across all agents i.e.
monolithic field effects. This is shown in Fig. \ref{emerge:FIELDFILL-MAG} and
\ref{emerge:FIELDFILL-DIST}, where the profiles of the two graphs are the same.
In addition, the new metric provides a mechanism for evaluating swarms that use
different algorithms to create a swarming effect and allows the algorithms'
inter-agent interactions to be compared rather than the resultant structures
that the algorithms create.
